Supplemental Material for ``Assessing United States county-level exposure for research on tropical cyclones and human health"

### 2020-02-24 writing/ehp_results.tex
Work on some final edits throughout

We explored potential limitations in these data by comparing them with data
from other available sources.  For estimates of storm-associated rainfall, we
compared estimates included in the open-source data with ground-based
observations in nine sample counties (Figure~\ref{fig:raincomparison}). Within
these counties, storm-related rainfall measurements were well-correlated
between the two data sources, with rank correlations (bottom right of each
graph in Figure~\ref{fig:raincomparison}) between 0.87 and 0.98. 
There was some evidence that our primary rainfall metric may tend to
underestimate rainfall totals in storms with extremely high rainfall, based on
a few heavy-rainfall storms in Harris County, TX, Mobile County, AL, Charleston
County, SC, and Wake County, NC (Figure~\ref{fig:raincomparison}). However,
even in these cases, it was rare for a storm to be classified differently
(under the precipitation threshold of 75~\si{\milli\metre} we used for binary
exposure classification for later analysis) based on the source of
precipitation data. Horizontal and vertical lines in each small plot in
Figure~\ref{fig:raincomparison} show the threshold of 75~\si{\milli\metre}, so
storms in the lower left and upper right quadrants would be classified the same
(``exposed'' or ``unexposed'') regardless of the precipitation data source,
while storms in the upper left and lower right quadrants would be classified
differently. Such cases were rare.

For peak sustained surface wind estimates, we found that the primary wind
exposure metric in the open-source data generally agreed well with the wind
radii reported in \ac{HURDAT2}. For most storms, $\ge$90\% of counties were
assigned the same category of wind speed ($<$34~kt; 34\,--\,49.9~kt;
50\,--\,63.9~kt; $\ge$64~kt) by both data sources
(Figure~\ref{fig:windcomparison}).  Disagreement was limited to a few storms
(e.g., Hurricanes Sandy in 2012 and Ike in 2008). For these two storms, the
modeled wind speed values used in the open-source data somewhat overestimated
the severity of the storm's winds at landfall but then underestimated the
extent of 34\,--\,49.9 kt winds from the storm's central track further inland
(Figure~S2). For epidemiology researchers who would like to conduct sensitivity
analysis between these two sets of wind data, we have included estimates from
the \ac{HURDAT2} wind radii as a secondary measure of county-level peak
sustained wind in the open-source dataset \parencite{hurricaneexposuredata}.

@@

When we calculated and mapped the average number of exposures per decade in
each county for single-hazard exposures (Figure~\ref{fig:averageexposure}),
strong geographical patterns were clear. Peak sustained wind exposure had a
strong coastal pattern, with almost all exposures in counties within about~200
\si{\kilo\metre} (124~mi) of the coastline. While tropical cyclone rain
exposure was also more frequent in coastal areas compared to inland areas,
there were also inland rain exposures in counties that were never classified as
exposed based on wind. Flood-based exposures were frequently in the
Mid-Atlantic region, with a pattern that skewed north compared with other
exposures. Rain and, to some extent, flood exposures were characterized by a
pattern defined by the Appalachian Mountains, with fewer exposures west of the
mountain range than to the east. Almost all tornado-based exposures were in
coastal states, with many in Florida, and almost none north of Maryland.
Patterns were similar when analysis was restricted to years with exposure data
for all four hazards available (1996\,--\,2011; Figure~S3).

@@

Finally, we assessed agreement between exposure classifications for each pair
of hazards, as well as with a proxy measurement based on
distance from the storm's track. We found that these exposure classifications
typically did not agree strongly between pairs of metrics---the set of counties
identified as ``exposed'' based on one metric often overlapped little with the set
identified as ``exposed'' by another metric, as storms frequently brought
different hazards to different locations.

@@

There were a few tropical cyclones in which exposure assessment agreed well
across several of the metrics considered.  For Floyd in~1999 (Figure~S4) and
Irene in~2011, for example, county-level classification agreed moderately to
well for all pairs of exposure metrics except those involving the tornado-based
metric (Figure~\ref{fig:jaccard}). For another set of tropical cyclones (e.g.,
Ernesto in~2006, and Bertha in~1996, Isabel in~2003), there was moderate to
good agreement for pairwise combinations of distance, rain, and wind, but poor
agreement for other combinations of metrics, while for another set of storms
(e.g., Matthew in~2004 and Katrina in~2005), there was moderate to good
agreement between distance and rain.

>>>>

We explored potential limitations in these data by comparing them with data
from other available sources.  For estimates of storm-associated rainfall, we
compared data in the open-source package with ground-based
observations in nine sample counties (Figure~\ref{fig:raincomparison}). Within
these counties, storm-related rainfall measurements were well-correlated
between the two data sources, with rank correlations (bottom right of each
graph in Figure~\ref{fig:raincomparison}) between 0.87 and 0.98. 
There was some evidence that our primary rainfall metric may tend to
underestimate rainfall totals in storms with extremely high rainfall, based on
a few heavy-rainfall storms in Harris County, TX, Mobile County, AL, Charleston
County, SC, and Wake County, NC (Figure~\ref{fig:raincomparison}). However,
even in these cases, it was rare for a storm to be classified differently
(under the precipitation threshold of 75~\si{\milli\metre} we used for binary
exposure classification for later analysis) based on the source of
precipitation data. Horizontal and vertical lines in each small plot in
Figure~\ref{fig:raincomparison} show the threshold of 75~\si{\milli\metre}, so
storms in the lower left and upper right quadrants would be classified the same
(``exposed'' or ``unexposed'') regardless of the precipitation data source,
while storms in the upper left and lower right quadrants would be classified
differently. Such cases were rare.

For peak sustained surface wind estimates, we found that the primary wind
exposure metric in the open-source data generally agreed well with the wind
radii reported in \ac{HURDAT2}. For most storms, $\>$90\% of counties were
assigned the same category of wind speed ($<$34~kt; 34\,--\,49.9~kt;
50\,--\,63.9~kt; $\ge$64~kt) by both data sources
(Figure~\ref{fig:windcomparison}).  Disagreement was limited to a few storms
(e.g., Hurricanes Sandy in 2012 and Ike in 2008). For these two storms, the
modeled wind speed in the open-source data somewhat overestimated the severity
of the storm's winds at landfall but then underestimated, particularly for
Hurricane Sandy, how far 34\,--\,49.9 kt winds extended from the storm's
central track further inland (Figure~S2). For epidemiology researchers who
would like to conduct sensitivity analysis using both sets of wind data,
we have included estimates from the \ac{HURDAT2} wind radii as a secondary
measure of county-level peak sustained wind in the open-source dataset
\parencite{hurricaneexposuredate}.

@@

When we calculated and mapped the average number of exposures per decade in
each county for single-hazard exposures (Figure~\ref{fig:averageexposure}),
strong geographical patterns were clear. Peak sustained wind exposure had a
strong coastal pattern, with almost all exposures in counties within about~200
\si{\kilo\metre} (124~mi) of the coastline. While tropical cyclone rain
exposure was also more frequent in coastal areas compared to inland areas,
there were also inland rain exposures in counties that were rarely or never
classified as exposed based on wind. Flood-based exposures were frequently in
the Mid-Atlantic region, with a pattern that skewed north compared with other
exposures. Rain and, to some extent, flood exposures were characterized by a
pattern defined by the Appalachian Mountains, with fewer exposures west of the
mountain range than to the east. Almost all tornado-based exposures were in
coastal states, with many in Florida, and almost none north of Maryland.
Patterns were similar when analysis was restricted to years with exposure data
for all four hazards available (1996\,--\,2011; Figure~S3).

@@

Finally, we assessed within-storm agreement between exposure classifications
for each pair of hazards, as well as with a proxy measurement based on distance
from the storm's track. We found that these exposure classifications typically
did not agree strongly between pairs of metrics---the set of counties
identified as ``exposed'' based on one metric often overlapped little with the
set identified as ``exposed'' by another metric, as storms frequently brought
different hazards to different locations.

@@

There were a few exceptions---tropical cyclones in which exposure assessment
agreed well across several of the metrics considered.  For Floyd in~1999
(Figure~S4) and Irene in~2011, for example, county-level classification agreed
moderately to well for all pairs of exposure metrics except those involving the
tornado-based metric (Figure~\ref{fig:jaccard}). For another set of tropical
cyclones (e.g., Ernesto in~2006, and Bertha in~1996, Isabel in~2003), there was
moderate to good agreement for pairwise combinations of distance, rain, and
wind, but poor agreement for other combinations of metrics, while for another
set of storms (e.g., Matthew in~2004 and Katrina in~2005), there was moderate
to good agreement between distance and rain.
